Federated Learning's Blessing:\\ FedAvg has Linear Speedup

### 2020-06-01 source/sec_SGD.tex
minor edits

is bounded, then $L$-smoothness implies $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$
for all $k$. In this overparameterized setting, we show in Section
\label{sec:overparameterized} that the convergence is in fact exponential,
which improves on the $\mathcal{O}(1/NT)$ rate in \cite{haddadpour2019convergence}
with better communication complexity.

>>>>

is bounded, then $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all
$k$. In this overparameterized setting, we show in Section~\ref{sec:overparameterized}
that the convergence is in fact exponential, thus improving on the
$\mathcal{O}(1/NT)$ rate in \cite{haddadpour2019convergence} with
better communication complexity.

### 2020-06-01 source/sec_problem.tex
minor edits

\textbf{Justfication for Assumptions~\ref{ass:boundedvariance} and~\ref{ass:subgrad2}.} These standard assumptions have been made in many previous
works in federated learning ~\cite{yu2019parallel,li2019convergence,stich2018local}. Here we provide one more justification
for their generality: even though they require bounds for all $\mathbf{w}$,
they are not much stronger than requiring bounds at the global optimal
parameter $\mathbf{w}^{\ast}$. As model average parameters become
closer to $\mathbf{w}^{\ast}$, the $L$-smoothness property implies
that $\mathbb{E}\|\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})\|^{2}$
and $\mathbb{E}\|\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})-\nabla F_{k}(\mathbf{w}_{t}^{k})\|^{2}$
will become close to $\mathbb{E}\|\nabla F_{k}(\mathbf{w}^{\ast},\xi_{t}^{k})\|^{2}$
and $\mathbb{E}\|\nabla F_{k}(\mathbf{w}^{\ast},\xi_{t}^{k})-\nabla F_{k}(\mathbf{w}^{\ast})\|^{2}$.
Therefore, there is no substantial difference between these assumptions
and assuming the bounds at $\mathbf{w}^{\ast}$ only. On the other
hand, some works such as \cite{haddadpour2019convergence} assume
bounds on the ``gradient diversity'', defined as $\frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$.
Such assumptions are more restrictive than our assumptions, because
they imply $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all local
objectives. For such ``overparameterized'' settings, we also prove
that the convergence is in fact exponential.

>>>>

\textbf{Justfication for Assumptions~\ref{ass:boundedvariance} and~\ref{ass:subgrad2}.} These standard assumptions have been made in many previous
works in federated learning ~\cite{yu2019parallel,li2019convergence,stich2018local}. Here we provide one more justification
for their generality: even though they require bounds for all $\mathbf{w}$,
they are not much stronger than requiring bounds at the global optimal
parameter $\mathbf{w}^{\ast}$. As model average parameters become
closer to $\mathbf{w}^{\ast}$, the $L$-smoothness property implies
that $\mathbb{E}\|\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})\|^{2}$
and $\mathbb{E}\|\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})-\nabla F_{k}(\mathbf{w}_{t}^{k})\|^{2}$
will become close to $\mathbb{E}\|\nabla F_{k}(\mathbf{w}^{\ast},\xi_{t}^{k})\|^{2}$
and $\mathbb{E}\|\nabla F_{k}(\mathbf{w}^{\ast},\xi_{t}^{k})-\nabla F_{k}(\mathbf{w}^{\ast})\|^{2}$.
Therefore, there is no substantial difference between these assumptions
and assuming the bounds at $\mathbf{w}^{\ast}$ only. On the other
hand, some works such as \cite{haddadpour2019convergence,li2018federated}
assume bounds on the ``gradient diversity'' or ``dissimilarity'',
defined as $\frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$.
Such assumptions are more restrictive than our assumptions, because
they imply $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all local
objectives. For such ``overparameterized'' settings, we will prove
that the convergence is in fact geometric.
